\chapter{Objectives}
\label{chap:objectives}

Following the enumeration made in \fullrefname{sec:context_contribution}, questions need to be asked in order to complete this project.
\begin{itemize}
    \item Which rules, best practices, technologies and aspects should be used in order to improve the security and confidentiality of online services?
    \item Which specific rules, best practices, technologies and aspects should be used in order to improve the security and confidentiality of the \gls{ai} field, particularly for \gls{ml} and \gls{dl} models?
    \item How can we provide an understandable and complete model of our findings?
\end{itemize}

The objectives below intent to provide answers to the questions asked.

% -----------------------------------------------------------------------------
\section{Primary Objectives}
\label{sec:objectives_primary}

All those objectives must be fulfilled in order to successfully complete the thesis.

\subsection{Establish an Up-To-Date Knowledge Collection}
\label{sec:objectives_primary_knowledge}

We need to collect an up-to-date and complete knowledge of the rules, best practices, technologies and aspects that contribute to enforce the security and privacy of online services. To this end, a proper literature review must be conducted. Its results will then be used as a foundation for the next phase of the thesis.

The content related to this objective must be included into the thesis report. It will list, explain and analyse the concepts found during the collection.

This objective must be completed before the end of the thesis, which is the \DTMdisplaydate{2023}{2}{10}{}.

\subsection{Provide an Understandable Guide or \Gls{framework}}
\label{sec:objectives_primary_guide}

Using the previously collected knowledge about security and privacy, a guide or \gls{framework} must be built. It must provide an understandable and appliable methodology in order to evaluate online services.

The content related to this objective must be included into the thesis report. It will list, explain and analyse the methodology used to build the guide or \gls{framework} and its content. The result of this objective will be an independent document, which must be attached to the thesis report as an appendix.

This objective must be completed before the end of the thesis, which is the \DTMdisplaydate{2023}{2}{10}{}.

\subsection{Apply and Test the Guide or \Gls{framework} on an Online Service}
\label{sec:objectives_primary_test}

Proper metrics must be used in order to conduct a proper evaluation of the guide or \gls{framework}. Once defined and explained, an evaluation of an online service including one or more \gls{ai} processes will be made, and the results will then be analysed and discussed.

The online service should be accessible and open, which means that we should have access to the source code. To this end, we could use the \citeproperref{Hestia}{https://bit.ly/3BzeDSN}{2022}{09}{22} ecosystem, which is an ongoing project leaded by \AdvisorOne.

The content related to this objective must be included into the thesis report. It will explain and analyse the results obtained during the evaluation.

This objective must be completed before the end of the thesis, which is the \DTMdisplaydate{2023}{2}{10}{}.

% -----------------------------------------------------------------------------
\section{Secondary Objectives}
\label{sec:objectives_secondary}

The secondary objective will only be completed if all primary objectives have been fulfilled and if there is still time left before the end of the project's timeframe.

\subsection{Submit the Guide as an Online Resource}
\label{sec:objectives_secondary_online}

Because of the nature of this thesis being an academical work and the ethics of such approaches, we believe that publishing the guide or \gls{framework} on the Internet would be useful to the public. It should therefore be published on some online platforms, such as a custom website, a forum or something equivalent.

The content related to this objective must be included into the thesis report. It will explain and analyse the approaches made in order to publish the guide or \gls{framework} online.

If this objective is considered, it must be completed before the end of the thesis, which is the \DTMdisplaydate{2023}{2}{10}{}.

% -----------------------------------------------------------------------------
\section{Constraints}
\label{sec:objectives_constraints}

Apart from the ones described in the respective objectives, an additional constraint must be respected regarding the whole thesis scope.

If any data collected under real-life conditions is used, analysed or processed during this thesis, ethical considerations and obligations must be applied.