\chapter{Activities}
\label{chap:activities}

The previously defined objectives imply several work activities in order to fulfil them. These activities are separated into more specific tasks. Please note that those tasks will not necessarily be done in a chronological order.

% -----------------------------------------------------------------------------
\section{Specifications}
\label{sec:activities_specifications}

This activity defines the scope and objectives of the thesis.

\begin{enumerate}
	\item Define the project specification
	\item Define the project planning
\end{enumerate}


% -----------------------------------------------------------------------------
\section{State of the Art}
\label{sec:activities_sota}

This activity covers the \fullrefname{sec:objectives_primary_knowledge} objective.

\begin{enumerate}
	\item Gather scientific articles and papers
	\item Gather online resources
	\item Gather standards
	\item Gather technologies
	\item Assess, analyse and explain the relevant resources
\end{enumerate}


% -----------------------------------------------------------------------------
\section{Methodology}
\label{sec:activities_methodology}

This activity partially covers the \fullrefname{sec:objectives_primary_guide} objective.

\begin{enumerate}
	\item Define and explain the guide or \gls{framework} chosen format
	\item Gather resources for the chosen format
	\item Prepare the document
\end{enumerate}

\newpage

% -----------------------------------------------------------------------------
\section{Build the Guide or \Gls{framework}}
\label{sec:activities_build}

This activity partially covers the \fullrefname{sec:objectives_primary_guide} objective.

\begin{enumerate}
	\item Define the categories
	\item Define the items to evaluate
	\item Define the evaluation process
\end{enumerate}


% -----------------------------------------------------------------------------
\section{Test and Evaluation}
\label{sec:activities_evaluation}

This activity covers the \fullrefname{sec:objectives_primary_test} objective.

\begin{enumerate}
	\item Define the proper metrics to evaluate the guide of \gls{framework}
	\item Get access to an online service
	\item Apply the methodology previously defined
	\item Evaluate and explain the results
\end{enumerate}

% -----------------------------------------------------------------------------
\section{Publish the Guide or \Gls{framework}}
\label{sec:activities_publish}

This activity covers the \fullrefname{sec:objectives_secondary_online} secondary objective.

\begin{enumerate}
	\item Assess the suitable online platforms
	\item Publish the document on selected platform(s).
\end{enumerate}


% -----------------------------------------------------------------------------
\section{Documentation}
\label{sec:activities_documentation}

The thesis report will be realized throughout the duration of the whole project, but this activity will be conducted at its end in order to change last details in said report. A poster must also be produced.

\begin{enumerate}
    \item Produce the thesis poster
    \item Finalize the thesis report
\end{enumerate}